# basic tidyr uses
`gather()` `spread()` can break down or transpose the data
the differnt between the two , 
is we use gather when the variable(s) is in the same property,
and use spread on a variable, which it's values are in the same property.

`separate()`for split the var with multiple property into adequate vars.


load in the files
```{r}
# getwd()
library(dplyr)
library(tidyr)
library(readr)
students <- read.table("./data/students.csv",sep = " ")
students2 <- read.table("./data/students2.csv",sep = " ")
students3 <- read.table("./data/students3.csv",sep = " ")
students4 <- read.table("./data/students4.csv",sep = " ")
sat <- read.table("./data/sat.csv",sep = " ")
```

# gather ()
gather() make the 2 variable, male and female, which cotains the count, into sex and count.
```{r}
head(students)
gather(students, sex, count, -grade)
# a <- gather(students, sex, count, -grade)
#spread(a, sex,count)
```
# gather() + separate()
students2 is a datsets contains the counts of 2 classes by sex,
what we want, is to separate the vars into class and sex with grade.
```{r}
head(students2)
```

first , we gather the datasets
```{r}
students2 %>%
  gather(sex ,count ,-grade )%>%
        head()
```
and then separate the column sex we've just made, into sex and class. 
```{r}

students2 %>%
  gather(sex ,count ,-grade ) %>%
  separate( col = sex, c("sex", "class")) %>%
  head()
```
# gather() + spread() + mutate()
the rows in students3 contains more than one information, 
we have to clean it up.
```{r}
head(students3)
```
first, gather up the classes
```{r}
students3 %>%
  gather(class, grade,-name,-test, na.rm = TRUE)%>%
        head()
```
spread function
```{r}
students3 %>%
  gather(class, grade, class1:class5, na.rm = TRUE) %>%
  spread(test, grade)%>%
        head()
```
add the col of  class by using `parse_number()`
```{r}
students3 %>%
  gather(class, grade, class1:class5, na.rm = TRUE) %>%
  spread(test, grade) %>%
        mutate(class = parse_number(class))%>%
  head
```
# break into 2 dataset by id
students4 is a dataset alike students3, but with id.
clening the data by using id as a key.
```{r}
head(students4)
```
       
```{r}
student_info <- students4 %>%
  select(id, name, sex) %>%
        unique()

gradebook <- students4 %>%
  select(id, class, midterm, final)

head(student_info)
head(gradebook)
```

# select() + gather() + separate() + group_by() + mutate()

```{r}
head(sat)
```

i want to know the differences between sex on sat subjects.
so i first select the var i need, 
which exclude the var contains total,
then gather the subject and sex(old column name),
into the new var named part_sex,
go on using separate to separate the part_sex var.
```{r}
sat %>%
  select(-contains("total")) %>%
        gather(key = part_sex, value = count, -score_range) %>%
        separate(col = part_sex,c("part", "sex"))%>%
        head
```

the goal is to know the difference between sex on the subjects, 
so we group up by sex and subjects (2*3)by using group_by function.
use mutate to make a apply of sum in all groups.
```{r}
sat %>%
  select(-contains("total")) %>%
  gather(part_sex, count, -score_range) %>%
  separate(part_sex, c("part", "sex")) %>%
 group_by(part, sex)%>%
  mutate(total = sum(count),
         prop = count/total
  ) %>% print
```

