---
title: "week 1"
author: "me"
date: "8/6/2021"
output: pdf_document
---
# week 1 notes for Cleaning Data

# dir
```{r}
if(!file.exists("data")){
        dir.create("data")
}
```

# Getting data from internet-download.file()
download.file() *url* *destfile* *method* 
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileUrl, destfile = "./data/2006.csv", methods = "curl")
list.files("./data")
```
dont't forget to add the access date.
```{r}
dateDownload <- date()
dateDownload
```
if the url start with **http** or **https** will be fine with using download.file()

# loading flat files - read.table()
read.table(*file*,*header*,*sep*,*row.names*,*nrows*)
*skip* *quote*
you'll get an error
```{r}
com <- read.table("./data/2006.csv")
head(com)

```

```{r}
com1 <- read.table("./data/2006.csv", sep = ",",header = TRUE, skip = 2, nrows = 8)
# com1
```

# 1 How many properties are worth $1,000,000 or more?
```{r}
com <- read.table("./data/2006.csv", sep = ",",header = TRUE)

a <- subset(com, VAL>=24)

# is.na(com$VAL)
```






# reading xlsx files
```{r}
if(!file.exists("data")){dir.create("data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
download.file(fileUrl, destfile = "./data/getdatagov.xlsx",method = "curl")
accessDate <- date()
```

# xlsx package
```{r}
#install.packages("xlsx")
library("xlsx")
gas <- read.xlsx("./data/getdatagov.xlsx",sheetIndex = 1, header = TRUE)
head(gas)
```



# 3 
Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:DAT
what's the value of:
sum(dat$Zip*dat$Ext,na.rm=T)?
```{r}
rowIndex <- 18:23
colIndex <- 7:15
dat <- read.xlsx("./data/getdatagov.xlsx",sheetIndex = 1 ,rowIndex = rowIndex,colIndex = colIndex, header = TRUE)
sum(dat$Zip*dat$Ext,na.rm=T)
```


# reading xml data
```{r}
# install.packages("XML")
library(XML)
file <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
restaurant <- xmlTreeParse(sub("s","",file),useInternalNode = TRUE)

#class(restaurant)
rootNode <- xmlRoot(restaurant)
xmlName(rootNode)
names(rootNode)
```


# 4
How many restaurants have zipcode 21231? 
```{r}
#rootNode[[1]][[3]][[2]]
zipcode <- xpathSApply(rootNode,"//zipcode",xmlValue)
# zipcode
d <- subset(zipcode,zipcode=="21231")

d <- subset(rootNode,xpathSApply(rootNode,"//zipcode",xmlValue)=="21231")
```

# data.table package
```{r}
#install.packages("data.able")
library("data.table")
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
c <- download.file(fileUrl, destfile = "./data/Idaho.csv")
idaho <- read.csv("./data/Idaho.csv",header = TRUE)

DT <- data.table(idaho)
# DT[1,]
# DT[c(3,5,10)]
DT[,list(mean(pwgtp1),sum(RAC1P))]
DT[,table(VPS)]
DT[,a:=RACBLK>0]
```


# 5
the ans is DT
cuz it uses the subset from packages of data tables
others uses bases packages
it'a a tricky one...
```{r}
# Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
# download.file(Url,destfile = "./data/Idaho.csv")
library(data.table)

Idaho <- read.csv("./data/Idaho.csv",header = TRUE)

Idaho <- data.table(Idaho)
file <- tempfile()
write.table(Idaho, file = file,row.names=FALSE,col.names = TRUE, sep = ",", quote = FALSE)
DT <- fread(file)

system.time(fread(file))
```

time <- 1000
a <- replicate(time,system.time(mean(DT[DT$SEX==1,]$pwgtp15),mean(DT[DT$SEX==2,]$pwgtp15)))

sum(a[1,]) #9.76
sum(a[3,]) #12.16
plot(b)

b <- replicate(time, system.time(DT[,mean(pwgtp15),by=SEX]))
sum(b[1,]) # 4.73
sum(b[3,]) #6.68

# error
# c <- replicate(time,system.time(rowMeans(DT)[DT$SEX==1],rowMeans(DT)[DT$SEX==2]))
#sum(c[1,])


# incorrect ans
#d <- replicate(time,system.time(mean(DT$pwgtp15,by=DT$SEX)))
#sum(d[1,]) #0.05

e <- replicate(time,system.time(sapply(split(DT$pwgtp15,DT$SEX),mean)))
sum(e[1,]) #0.47
sum(e[3,]) #0.49

f <- replicate(time,system.time(tapply(DT$pwgtp15,DT$SEX,mean)))
sum(f[1,]) #0.52
sum(f[3,]) #0.55
